---
title: "Portfolio Assignment 3: Leaflet Maps"
author: "Liv Lindeneg"
date: "26/05-2026"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(tidyverse)
library(leaflet)
library(htmlwidgets)
library(googlesheets4)
```

## Making Maps with Leaflet

- This assignment uses and explores Leaflet to make maps of Denmark.
- First a more basic map of Denmark with different layers will be made.
- Afterwards that map will be modified with specific points based on data made by fellow students in class.
- This assignment will also look at different ways of visualizing the points and how to give the points popups.

### Task 1: Create a Danish equivalent of AUSmap with Esri layers 

- The method to making the map of Denmark follows the tutorial we were presented with in class of an Australian map

```{r Denmark Map, eval=TRUE}
# Coordinates: 56.1521976,10.1974321 from Google Maps

l_dan <- leaflet() %>%
  setView(10.1974321, 56.1521976, zoom = 13)

esri <- grep("^Esri", providers, value = TRUE)

for (provider in esri) {
  l_dan <- l_dan %>% addProviderTiles(provider, group = provider)
}

DANmap <- l_dan %>%
  addLayersControl(baseGroups = names(esri),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  addMiniMap(tiles = esri[[1]], toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
  htmlwidgets::onRender("
                        function(el, x) {
                        var myMap = this;
                        myMap.on('baselayerchange',
                        function (e) {
                        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                        })
                        }") %>% 
  addControl("", position = "topright")

DANmap

saveWidget(DANmap, "DANmap.html", selfcontained = TRUE)

```

### Task 2: Read in the googlesheet data you and your colleagues created into your DANmap object

- Using the data from the googlesheet created in class and add it to the Denmark map created in task 1.
- The data is over favorite places in Denmark.
- This will make a map with specific location points.

```{r Points from googlesheet, eval=TRUE, message=FALSE, warning=FALSE}

gs4_deauth()

places <- read_sheet("https://docs.google.com/spreadsheets/d/1PlxsPElZML8LZKyXbqdAYeQCDIvDps2McZx1cTVWSzI/edit#gid=124710918", col_types = "cccnncnc", range = "DAM2026")

glimpse(places) 

DANmap2 <- DANmap %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = paste(places$Placename, "<br>", places$Type))

DANmap2

saveWidget(DANmap2, "Prtf3_PointMap.html", selfcontained = TRUE)  

```

### Task 3: Can you cluster the points in Leaflet?

- Another way of visualizing the location points are by clustering the data on the map.
- This will group the data together in areas where there are multiple data points.

```{r Clustering, eval=TRUE, warning=FALSE}

DANmap3 <- DANmap %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = paste(places$Placename, "<br>", places$Type),
             clusterOptions = markerClusterOptions())

DANmap3

saveWidget(DANmap3, "Prtf3_ClusterMap.html", selfcontained = TRUE)  

```



### Task 4: Look at the two maps (with and without clustering) and consider what each is good for and what not.

- When you cluster the data as in task 3, it makes a map that makes it very easy to see if there are any trends geographically over the places picked by students. It look very clean and simple.
- When using the points like in task 2, you can see specifically where these locations or places are. It can look a bit messy especially if there are a lot of data on the map, but it maps the data more precisely. When you cluster the data you have to zoom in to the individual data points. 



### Task 5: Find out how to display the notes and classifications column in the map. 

- The previous maps points have had the "placename" and "type" displayed in the popup.
- We can add more columns in the popups if we want.
- Have here added the "Description" column as well.

```{r Popups, eval=TRUE, warning=FALSE}

DANmap4 <- DANmap %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = paste(places$Placename, "<br>", places$Type, "<br>", places$Description))

DANmap4

saveWidget(DANmap4, "Prtf3_PopupExpanded.html", selfcontained = TRUE)  

```



