---
title: "Portfolio Assignment 5: Women, Marriage and Employment"
author: "Liv Lindeneg"
date: "26/05-2026"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(tidyverse)
ugifte_kvinder <- read_csv("data/kvinder_ugifte_1860.csv")
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

## Making a top 10 for all unmarried women 1860 Aarhus

```{r Top 10 occupations}

# Making a vector over descriptions that needs to be excluded
# Not occupations
ikke_erhverv <- c("datter", "plejedatter", "barn", "søn", "kone", "søster", "fattiglem", "huusfader", "indsidder",
                  "forsørges af familie", "barnebarn", "steddatter", "tjenestekarl", "familiemedlem", "tjener forældrene",
                  "dagleier", "gaardmand", "huusmand")


# Making a top 10 with only occupations
# Keep adding to the vector above until the top 10 is only occupations
top_erhverv <- ugifte_kvinder %>% 
  filter(!stilling %in% ikke_erhverv) %>%
  count(stilling, sort = TRUE) %>%
  slice_head(n = 10)

top_erhverv

# Mutate to get a column showing percentage as well
percentage <- ugifte_kvinder %>%
  count(stilling, sort = TRUE) %>% 
  mutate(percentage = n/sum(n)*100) %>% 
  filter(!stilling %in% ikke_erhverv) %>%
  slice_head(n = 10)

percentage
```


## Most common occupations for all unmarried women 1860 Aarhus

#### The amount in numbers

```{r most common occupations}
# Making a plot
# Showing the amount in numbers
plot_toperhverv <- top_erhverv %>% 
  ggplot(aes(x = reorder(stilling,n), y=n))+
  geom_col(fill = "salmon") +
  geom_text(aes(label = n),
            hjust = 0.5,
            size = 3) +
  coord_flip() +labs(
    title = "Top 10 occupations for all unmarried women 1860 Aarhus",
    x = "Occupation",
    y = "Count"
  ) +
  theme_bw()+
  theme(text=element_text(size=12))

plot_toperhverv
```



#### The amount in percentage

```{r most common occupations percentage}

# Making a plot
# Showing the amount in percentage
percentage_toperhverv <- percentage %>% 
  ggplot(aes(x = reorder(stilling, percentage), y=percentage))+
  geom_col(fill = "palegreen") +
  geom_text(aes(label = paste0(round(percentage, 2), "%")),
            hjust = 0.5,
            size = 3) +
  coord_flip() +labs(
    title = "Top 10 occupations for all unmarried women 1860 Aarhus",
    x = "Occupation",
    y = "Percentage"
  ) +
  theme_bw()+
  theme(text=element_text(size=12))

percentage_toperhverv
```

## Making a top 10 for unmarried women over 25 years old 1860 Aarhus

```{r top 10 occupations}
# Filter to get only the women that are older than 25
stilling_over25 <- ugifte_kvinder %>% 
  filter(alder > 25) %>% 
  count(stilling, sort = TRUE)


# Making a top 10 for occupations for women over 25
# Excluding non occupation descriptions 
top_over25 <- stilling_over25 %>% 
  filter(!stilling %in% ikke_erhverv) %>%
  slice_head(n = 10)

top_over25

# Adding a column with percentage
# Over 25 but percentage instead
percentage_over25 <- stilling_over25 %>% 
  mutate(percentage = n/sum(n)*100) %>% 
  filter(!stilling %in% ikke_erhverv) %>%
  slice_head(n = 10)

percentage_over25
```



## Most common occupations for unmarried women over 25 years old 1860 Aarhus

#### The amount in numbers

```{r most common occupations over 25}

# Making a plot
# Showing the amount in numbers
plot_over25 <- top_over25 %>% 
  ggplot(aes(x = reorder(stilling,n), y=n))+
  geom_col(fill = "lightskyblue1") +
  geom_text(aes(label = n),
            hjust = 0.5,
            size = 3) +
  coord_flip() +labs(
    title = "Top 10 occupations for unmarried women over 25 1860 Aarhus",
    x = "Occupation",
    y = "Count"
  ) +
  theme_bw()+
  theme(text=element_text(size=12))

plot_over25

```

#### The amount in percentage

```{r most common occupations over 25 percentage}

# Making a plot
# Showing the amount in percentage
plot_over25percentage <- percentage_over25 %>% 
  ggplot(aes(x = reorder(stilling, percentage), y=percentage))+
  geom_col(fill = "plum2") +
  geom_text(aes(label = paste0(round(percentage, 2), "%")),
            hjust = 0.5,
            size = 3) +
  coord_flip() +labs(
    title = "Top 10 occupations for unmarried women over 25 1860 Aarhus",
    x = "Occupation",
    y = "Percentage"
  ) +
  theme_bw()+
  theme(text=element_text(size=12))

plot_over25percentage

```



## Age distibution for the most common occupations

#### Filtering

```{r filtering}

# Making a vector of the most common occupation
# The vector is needed for filtering
top_vector <- top_erhverv %>% 
  pull(stilling) # to make a vector in order to filter

# Filtering all the occupations and corresponding ages for the most common occupations
alder_stilling <- ugifte_kvinder %>% 
  select(alder, stilling) %>% 
  filter(stilling %in% top_vector)
```


#### Plot
```{r age distribution, warning=FALSE}

# Making a plot
# Multiple histograms using facets_wrap()
plot_agedistri <- alder_stilling %>% 
  ggplot(aes(x=alder, fill=stilling))+
  geom_histogram(bins=20, color="white")+
  facet_wrap(~stilling, scales="free_y")+
  labs(title="Age distribution for the top 10 occupations among unmarried women 1860 Aarhus",
       x="Age",
       y="Count")+
  theme_bw()+
  theme(text=element_text(size=10))+
  theme(legend.position = "none")

plot_agedistri
```
